---
title: An R Markdown document converted from "main.ipynb"
output: html_document
---

##### Setup

```{r}
suppressPackageStartupMessages(library("NbClust"))
suppressPackageStartupMessages(library(hopkins))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library("ggpubr"))
suppressPackageStartupMessages(library("stats"))
suppressPackageStartupMessages(library("FactoMineR"))
suppressPackageStartupMessages(library("factoextra"))
suppressPackageStartupMessages(library(corrplot))   
suppressPackageStartupMessages(library(tidyverse))
library(factoextra)
library(fpc)
library(NbClust)
library("ggcorrplot")
library(pheatmap)
suppressPackageStartupMessages(library(gridExtra))
library(RColorBrewer)
library(ggridges)
```

```{r}
df <- read.csv('Country-data.csv', stringsAsFactors = T)
dfDict <- read.csv('data-dictionary.csv', stringsAsFactors = T)

cont <- read.csv('countries_info.csv', stringsAsFactors = F)
col <- c("name","Continent")
cont <- cont[, col]

dfWithCont <- merge(df, cont, by.x="country", by.y="name",all.x = TRUE)
dfWithCont <- dfWithCont %>% remove_rownames %>% column_to_rownames(var="country")

df <- df %>% remove_rownames %>% column_to_rownames(var="country")
col_order <- c("gdpp", "income", "inflation",
               "exports", "imports","health","life_expec","child_mort","total_fer")
df <- df[, col_order]
```

```{r}
dfWithCont$Continent[5] <- "NAs"
dfWithCont$Continent[11] <- "NAs"
dfWithCont$Continent[14] <- "NAs"
dfWithCont$Continent[17] <- "NAs"
dfWithCont$Continent[21] <- "EU"
dfWithCont$Continent[30] <- "NAs"
dfWithCont$Continent[38] <- "AF"
dfWithCont$Continent[39] <- "NAs"
dfWithCont$Continent[40] <- "NAs"
dfWithCont$Continent[41] <- "AF"
dfWithCont$Continent[46] <- "NAs"
dfWithCont$Continent[49] <- "NAs"
dfWithCont$Continent[62] <- "NAs"
dfWithCont$Continent[63] <- "NAs"
dfWithCont$Continent[67] <- "NAs"
dfWithCont$Continent[77] <- "NAs"
dfWithCont$Continent[84] <- "AS"
dfWithCont$Continent[85] <- "AS"
dfWithCont$Continent[93] <- "EU"
dfWithCont$Continent[102] <- "OC"
dfWithCont$Continent[118] <- "NAs"
dfWithCont$Continent[135] <- "EU"
dfWithCont$Continent[142] <- "NAs"
dfWithCont$Continent[159] <- "EU"
dfWithCont$Continent[160] <- "NAs"
```

```{r}
dfScale <- scale(df)
```

##### Exploration

```{r}
options(repr.plot.width=10, repr.plot.height=6)
boxplot(df)
```

```{r}
corr <- round(cor(df), 1)
ggcorrplot(corr, p.mat = cor_pmat(df),
hc.order = TRUE, type = "lower",
color = c("#00AFBB", "white", "#FC4E07"),
outline.col = "white", lab = TRUE)
```

##### PCA Exploration

```{r}
res.pca <- PCA(dfWithCont[,-10], graph = FALSE,scale.unit=TRUE)
```

```{r}
# An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the
# original variables in standardized data. This is commonly used as a cutoff point for which PCs
# are retained. This holds true only when the data are standardized.

eig.val <- get_eigenvalue(res.pca)
eig.val
```

```{r}
options(repr.plot.width=10, repr.plot.height=6)
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```

```{r}
res.pca$var$coord[,1:2]
```

```{r}
pc12 <- as.data.frame(res.pca$var$coord[,1:2])
```

```{r}
fviz_pca_var(res.pca, col.var = "cos2",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE)
```

```{r}
fviz_cos2(res.pca, choice = "var", axes = 1:2)
```

```{r}
corrplot(res.pca$var$contrib, is.corr=FALSE)
```

```{r}
# The red dashed line on the graph above indicates the expected average contribution. If the contribution
# of the variables were uniform, the expected value would be 1/length(variables) = 1/10 = 10%. For a
# given component, a variable with a contribution larger than this cutoff could be considered as important
# in contributing to the component.

fviz_contrib(res.pca, choice = "var", axes = 1:2, top = 10)
```

```{r}
options(repr.plot.width=10, repr.plot.height=10)
fviz_pca_biplot(
res.pca,
col.ind = dfWithCont$Continent, 
palette = c("#00AFBB", "#E7B800", "#FC4E07","#44AC44","#463fca","#d152cc"),
label = "var",
col.var = "black", 
repel = TRUE,
legend.title = "Continents",
pointshape=19,
alpha.var=0.4,
pointsize = 3,
)
```

```{r}
fviz_pca_biplot(res.pca,repel=TRUE)
```

#### Clustering exploration

##### mm

```{r}
set.seed(123)
hopkins(scale(dfScale))
hopkins.pval(0.999995046082251, dim(dfScale)[1])
```

```{r}
res.dist <- get_dist(scale(df), method = "pearson")
fviz_dist(res.dist, lab_size = 8,show_labels = FALSE)
```

##### K-Means

```{r}
options(repr.plot.width=10, repr.plot.height=6)
fviz_nbclust(scale(df), kmeans, method = "wss") +
labs(subtitle = "Elbow method")
fviz_nbclust(scale(df), kmeans, method = "silhouette")+
labs(subtitle = "Silhouette method")
```

If we want to identify clusters of observations with the same overall profiles regardless
of their magnitudes, then we should go with correlation-based distance as a dissimilarity
measure. This is particularly the case in gene expression data analysis, where we
might want to consider genes similar when they are “up” and “down” together. It is
also the case, in marketing if we want to identify group of shoppers with the same
preference in term of items, regardless of the volume of items they bought.
If Euclidean distance is chosen, then observations with high values of features will be
clustered together. The same holds true for observations with low values of features.

```{r}
k2 <- kmeans(scale(df), centers = 2, nstart = 50)
k3 <- kmeans(scale(df), centers = 3, nstart = 50)
k4 <- kmeans(scale(df), centers = 4, nstart = 50)
k5 <- kmeans(scale(df), centers = 5, nstart = 50)
k6 <- kmeans(scale(df), centers = 6, nstart = 50)
k7 <- kmeans(scale(df), centers = 7, nstart = 50)

# plots to compare
p1 <- fviz_cluster(k2, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 2),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 3),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 4),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 5),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 5")
p5 <- fviz_cluster(k6, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 6),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 6")
p6 <- fviz_cluster(k7, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 7),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 7")

options(repr.plot.width=20, repr.plot.height=20)
grid.arrange(p1, p2, p3, p4,p5,p6,nrow = 3)
```

```{r}
options(repr.plot.width=10, repr.plot.height=8)
fviz_cluster(k6, data = scale(df),
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 6),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal(),
) + geom_point(data=pc12,aes(x=Dim.1*8,y=Dim.2*8), size=0.1) +
  geom_segment(data=pc12,aes(x=0, y=0, xend=Dim.1*8, yend=Dim.2*8),arrow = arrow(length = unit(0.2, "cm"))) + 
  geom_text(data=pc12,aes(x=Dim.1*8,y=Dim.2*8),label = rownames(pc12)) 
```

##### PAM

```{r}
options(repr.plot.width=10, repr.plot.height=6)
fviz_nbclust(dfScale, pam, method = "silhouette")+
theme_classic()
# Elbow method
fviz_nbclust(dfScale, pam, method = "wss") +
labs(subtitle = "Elbow method")
```

```{r}
pam2 <- pam(dfScale, 2,metric="manhattan")
pam3 <- pam(dfScale, 3,metric="manhattan")
pam4 <- pam(dfScale, 4,metric="manhattan")
pam5 <- pam(dfScale, 5,metric="manhattan")
pam6 <- pam(dfScale, 6,metric="manhattan")
pam7 <- pam(dfScale, 7,metric="manhattan")

# plots to compare
p1 <- fviz_cluster(pam2,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 2),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 2")
p2 <- fviz_cluster(pam3,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 3),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 3")
p3 <- fviz_cluster(pam4,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 4),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 4")
p4 <- fviz_cluster(pam5,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 5),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 5")
p5 <- fviz_cluster(pam6,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 6),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 6")
p6 <- fviz_cluster(pam7,
palette = get_palette(c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), 7),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
) + ggtitle("k = 7")

options(repr.plot.width=20, repr.plot.height=20)
grid.arrange(p1, p2, p3, p4,p5,p6,nrow = 3)
```

##### Exploration of k=3

```{r}
options(repr.plot.width=15, repr.plot.height=6)

scaleMedoids<- dfScale[row.names(pam3$medoids),]
breaksList = seq(-0.80, 0.80, by = 0.05)
pheatmap(scaleMedoids,display_numbers = TRUE,
         number_color = "black", 
         fontsize=15,
         fontsize_number = 13,cluster_cols = FALSE,
         cluster_rows = FALSE,cellwidth=70,cellheight=70,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "OrRd")))(length(breaksList)), 
         breaks = breaksList)

dd3 <- cbind(df, cluster = pam3$cluster)
dd3[row.names(pam3$medoids),]
```

```{r}
saved <- dd3[row.names(pam3$medoids),]
saved <- as.data.frame(saved)
saved <- data.frame(saved[,-10], row.names = saved[,10])
saved <- as_tibble(cbind(nms = names(saved), t(saved)))

saved <- as.data.frame(saved)
saved[] <- lapply(saved, function(x) as.numeric(as.character(x)))
saved$nms <- c("gdpp", "income", "inflation",
               "exports", "imports","health","life_expec","child_mort","total_fer")

colnames(saved) <- c("cols","cl1","cl2","cl3")
saved <- saved %>% mutate(cols = factor(cols, levels = unique(cols)))
```

```{r}
options(repr.plot.width=8, repr.plot.height=6)

saved %>% 
  ggplot(aes(x=cols, y=cl1,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 1 - Undeveloped Countries - Tanzania") +ylab("")+xlab("") +
  theme(axis.text = element_text(size = 12))  

saved %>% 
  ggplot(aes(x=cols, y=cl2,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 2 - Developing Countries - Tunisia") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))  

saved %>% 
  ggplot(aes(x=cols, y=cl3,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 3 - Developed Countries - Finland") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))  
```

```{r}
options(repr.plot.width=7, repr.plot.height=6)
dd3 <- cbind(dfWithCont, cluster = pam3$cluster)

dd3 %>%
group_by(cluster,Continent) %>% tally() %>%
ggplot(aes(x=cluster, y=n, fill=Continent)) +
scale_fill_manual(values=get_palette(c("#2E9FDF", "#E7B800", "#FC4E07"), 6)) +
geom_bar(stat="identity", position=position_dodge()) + 
xlab("Cluster") +
ylab("")+
  theme(axis.text = element_text(size = 12))  
```

##### Exploration of k=5

```{r}
dd5 <- cbind(dfWithCont, cluster = pam5$cluster)
dd5scale <- cbind(dfScale, cluster = pam5$cluster)
```

```{r}
res.pca <- PCA(dfScale, graph = FALSE)
```

```{r}
# plots to compare
p1 <- fviz_pca_biplot(res.pca, select.ind = list(name=row.names(dd5 %>%
filter(cluster==1))),repel=TRUE) +
ggtitle("Cluster 1")
p2 <- fviz_pca_biplot(res.pca, select.ind = list(name=row.names(dd5 %>%
filter(cluster==2))),repel=TRUE) + ggtitle("Cluster 2")
p3 <- fviz_pca_biplot(res.pca, select.ind = list(name=row.names(dd5 %>%
filter(cluster==3))),repel=TRUE) + ggtitle("Cluster 3")
p4 <- fviz_pca_biplot(res.pca, select.ind = list(name=row.names(dd5 %>%
filter(cluster==4))),repel=TRUE) + ggtitle("Cluster 4")
p5 <- fviz_pca_biplot(res.pca, select.ind = list(name=row.names(dd5 %>%
filter(cluster==5))),repel=TRUE) + ggtitle("Cluster 5")

options(repr.plot.width=20, repr.plot.height=20)
grid.arrange(p1, p2, p3, p4,p5,nrow = 3)
```

```{r}
options(repr.plot.width=15, repr.plot.height=6)

scaleMedoids<- dfScale[row.names(pam5$medoids),]
breaksList = seq(-0.80, 0.80, by = 0.05)
pheatmap(scaleMedoids,display_numbers = TRUE,
         number_color = "black", 
         fontsize=15,
         fontsize_number = 13,cluster_cols = FALSE,
         cluster_rows = FALSE,cellwidth=70,cellheight=70,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "OrRd")))(length(breaksList)), 
         breaks = breaksList)

dd5[row.names(pam5$medoids),]
```

```{r}
saved <- as.data.frame(dd5scale) %>%
group_by(cluster) %>% summarise_all(list(median))
saved <- as.data.frame(saved)
saved <- data.frame(saved[,-1], row.names = saved[,1])
saved <- as_tibble(cbind(nms = names(saved), t(saved)))

saved <- as.data.frame(saved)
saved[] <- lapply(saved, function(x) as.numeric(as.character(x)))
saved$nms <- c("gdpp", "income", "inflation",
               "exports", "imports","health","life_expec","child_mort","total_fer")

colnames(saved) <- c("cols","cl1","cl2","cl3","cl4","cl5")
saved <- saved %>% mutate(cols = factor(cols, levels = unique(cols)))
```

```{r}
options(repr.plot.width=8, repr.plot.height=6)

p1 <- saved %>% 
  ggplot(aes(x=cols, y=cl1,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 1 - 1 - Tanzania") +ylab("")+xlab("") +
  theme(axis.text = element_text(size = 12))  

p2 <- saved %>% 
  ggplot(aes(x=cols, y=cl2,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 2 - 3 - Albania") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))  

p3<- saved %>% 
  ggplot(aes(x=cols, y=cl3,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 3 - 2 - Bolivia") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))

p4 <- saved %>% 
  ggplot(aes(x=cols, y=cl4,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 4 - 4 - Estonia") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))  

p5 <- saved %>% 
  ggplot(aes(x=cols, y=cl5,fill=cols)) + 
  geom_bar (stat="identity") + ylim(-1.85,1.85) + guides(fill = "none")+
  ggtitle("Cluster 5 - 5 - Finlandia") +ylab("")+xlab("")+
  theme(axis.text = element_text(size = 12))

options(repr.plot.width=20, repr.plot.height=20)
grid.arrange(p1, p2, p3, p4,p5,nrow = 3)
```

